---
title: "p8105_hw3_yw4251"
author: "You Wu"
date: 2023-10-10
output: github_document
---
# Problem 0
```{r}
library(tidyverse)
library(p8105.datasets)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .8,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

First, load the data.

```{r}
data("instacart")
instacart
```

## Description
This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, where each row in the dataset is a product from an order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.


## How many aisles are there, and which aisles are the most items ordered from?
```{r}
aisle_counts=
  instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n_obs=n()
  ) %>% 
  arrange(-n_obs)
```

There are `r nrow(aisle_counts)` aistles, and `r aisle_counts[1,1]` are the most items ordered from.

## Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it. As shown:

Aisles are ordered by ascending number of items. The majority of aisles have a low order count, close to the 10,000 mark or below. However, as we move to the right side of the graph, there is a noticeable increase in the number of items ordered for `fresh vegetables` and `fresh fruits`.

```{r}
aisle_counts %>% 
  filter(n_obs>10000) %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n_obs)) %>% 
  ggplot(aes(x=aisle,y=n_obs))+
  geom_point()+
  labs(
    title = "Number of items ordered in each aisle",
    x = "Aisle Names",
    y = "Item Counts in Each Aisle",
    caption = "Data from the instacart")+
  scale_y_continuous(
    breaks = c(10000, 60000, 110000, 160000),
    labels = c("10000","60000", "110000","160000"))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1),
        plot.title = element_text(hjust = 0.5))
```

## Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. As shown:

`Baking Ingredients Aisle`: Sweeteners like "Light Brown Sugar" and leavening agents such as "Pure Baking Soda" and "Cane Sugar" are popular choices.

`Dog Food Care Aisle`: Premium dog treats and high-quality dog food like "Organic Chicken & Brown Rice Recipe" are favored.

`Packaged Vegetables Fruits Aisle`: Organic and healthy options like "Organic Baby Spinach" and "organic berries" are in high demand.

```{r}
instacart %>% 
  filter(aisle%in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle,product_name) %>% 
  summarize(n_obs=n()) %>% 
  arrange(-n_obs) %>% 
  slice_head(n=3) %>% 
  mutate(rank = min_rank(desc(n_obs))) %>% 
  knitr::kable()
```

## Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. As shown:

Pink Lady Apples tend to be purchased a bit earlier in the day compared to Coffee Ice Cream, except for day 5

```{r}
instacart %>% 
  filter(product_name%in%c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name,order_dow) %>% 
  summarize(mean_hour=mean(order_hour_of_day)) %>% 
  pivot_wider(names_from = order_dow,
              values_from=mean_hour) %>% 
  knitr::kable(digits=2)
```


# Problem 2
```{r}
data("brfss_smart2010")
```

First, data cleaning. I standardize the column names, and filters rows where the topic is "Overall Health". It then reorders the responses in the 'response' column into an ascending order of health quality. The resulting cleaned data is stored in `cleaned_brfss`. 

```{r}
cleaned_brfss=
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic=="Overall Health") %>% 
  mutate(response=factor(
    response,
    levels = c("Poor", "Fair", "Good", "Very good", "Excellent"),
    ordered = TRUE))
```

The dataset includes `r nrow(cleaned_brfss)` rows and `r ncol(cleaned_brfss)` columns. Important variables include class, topic, data value, response level, etc.

## In 2002, which states were observed at 7 or more locations? What about in 2010? As shown:

In 2002, CT, FL, MA, NC, NJ, PA were observed at 7 or more locations; In 2010, CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA were observed at 7 or more locations.

```{r}
cleaned_brfss %>% 
  filter(year%in%c(2002,2010)) %>%
  group_by(year,locationabbr) %>% 
  summarize(n_loc=n_distinct(locationdesc)) %>% 
  filter(n_loc>=7) %>% 
  select(-n_loc) %>% 
  pivot_wider(
    names_from=year,
    values_from=locationabbr
  ) %>% 
  knitr::kable()
```

## Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help). As shown:

The Spaghetti Plot displays the average values from 2002 to 2010 for various U.S. states. It shows significant variability across states. A noticeable trend is the dip in average values for many states around 2006, followed by a subsequent recovery. The values range between approximately 15 and 30. There's no uniform trend evident for all the states across the shown years.

```{r}
excel_cleaned_brfss=
  cleaned_brfss%>% 
  filter(response=="Excellent") %>% 
  group_by(year,locationabbr) %>% 
  summarize(avg_data_val=mean(data_value))

ggplot(excel_cleaned_brfss,aes(x=year,y=avg_data_val,color=locationabbr))+
  geom_line(size = 0.8,alpha=0.7)+
  theme_classic()+
  labs(
    title = "Spaghetti Chart of Average Value over Time by States",
    x = "Year",
    y = "Average Data Values Across locations within State") +
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.text = element_text(size = 8),
    legend.position = "right",
    legend.key.size = unit(0.5, "cm")
  )
  
```

## Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State. As shown:

The distributions for the categories "Very good" and "Good" remain relatively stable across both years, indicating a consistent perception within these health categories. However, a more granular examination reveals a subtle decline in the "Excellent" category by 2010 and a concomitant increase in the "Poor" health category. These shifts, though not drastic, suggest potential underlying changes in the health landscape or in the subjective health evaluations of New Yorkers during this period. 


```{r}
ny_cleaned_brfss=
  cleaned_brfss %>% 
  filter(year%in%c(2006,2010)&locationabbr=="NY")


ggplot(ny_cleaned_brfss,aes(x=data_value,y=response,fill=response))+
  geom_density_ridges(alpha=0.3,scale=0.7,bins=40)+
  facet_grid(~year)+
  labs(
    title = "Distribution of Response Values in NY State (2006 vs 2010)",
    x = "Data Value",
    y = "Ordered Response Level",
    caption = "Data from the NY Cleaned Brfss"
    ) +
  theme_ridges(font_size = 14)+
  theme(
    legend.position = "none",
    axis.title.x = element_text(hjust = 0.5),
    axis.title.y = element_text(hjust = 0.5)
  )

```

I also make a box plot to better investigate the distribution value change. Both years show higher median values for "Good" and "Very good" responses. The "Good" and "Very good" categories exhibit higher median values compared to other health assessments. The "Excellent" category sees an upward shift in 2010, while "Fair" remains stable across both years. Notably, the "Poor" category has the lowest median values in both years, with minimal variability. Overall, positive health perceptions prevail, with minor changes between the two years.
```{r}
ggplot(ny_cleaned_brfss,aes(x=response,y=data_value,fill=response))+
  geom_boxplot()+
  facet_wrap(~year)+
  labs(title="Distribution of Response Values in NY State (2006 vs 2010)",
       x="Response", 
       y="Data Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


# Problem 3
First, clean and merge the dataset. 
```{r}
covar_df=read_csv("data/nhanes_covar.csv",skip=4) %>% 
  janitor::clean_names() %>% 
  drop_na() %>%
  filter(age>=21) %>% 
  mutate(
    sex=as.factor(sex),
    sex=recode(sex, "1"="male", "2"="female"),
    education=factor(
      education,
      levels = c("1", "2", "3"),
      labels = c("Less than high school", "High school equivalent", "More than high school"),
      ordered = TRUE)
         )
  
accel_df=read_csv("data/nhanes_accel.csv") %>% 
  janitor::clean_names()

covar_add_accel=left_join(covar_df,accel_df,by="seqn") %>% 
  drop_na()
  
```
The merged dataset includes `r nrow(covar_add_accel)` rows and `r ncol(covar_add_accel)` columns, where each row represent an participant. Variables include sex, age, education, bmi and MIMS values for each minute.

## Produce a reader-friendly table for the number of men and women in each education category.

The table shows the distribution of men and women across education levels. Both genders are almost even in the "Less than high school" and "More than high school" categories. However, there are more males (35) than females (23) in the "High school equivalent" category.

```{r}
covar_add_accel %>% 
  group_by(education,sex) %>% 
  summarize(count=n()) %>% 
  knitr::kable()
```
## Create a visualization of the age distributions for men and women in each education category. Comment on these items. As shown:

For those with "Less than high school", median ages are similar for both genders. In the "High school equivalent" category, males have a higher median age than females. However, in the "More than high school" group, both genders exhibit closely aligned median ages, with females having a slightly broader distribution.

```{r}
ggplot(covar_add_accel,aes(x=sex,y=age,fill=sex))+
  geom_boxplot(alpha=0.5)+
  facet_grid(~education) + 
  labs(
    title = "Age Distributions by Sex Across Education Categories",
    x = "Time",
    y = "Age",
    fill = "Sex"
    )+
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

## Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot. As shown:

There's a noticeable decline in activity with age across all education levels.For individuals with "Less than high school" education, the decline is more pronounced, especially among males. Those with a "High school equivalent" education display a less steep decline, but males still show a greater reduction than females. Among the "More than high school" educated group, their decline shows least steep, and both genders show similar activity decline rates. Across all categories, males generally have higher activity levels at younger ages compared to females.


```{r}
covar_add_accel %>% 
  mutate(total_activity=rowSums(select(., min1:min1440))) %>%
  ggplot(aes(x=age,y=total_activity,color=sex))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  facet_grid(~education)+
   labs(
    title = "Total Activity by Age, Sex, and Education Level",
    x = "Age",
    y = "Total Activity",
    color = "Sex"
  )+
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```


## Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences. As shown:

The 24-hour activity graph showcases distinct patterns based on education level and gender. Across all educational backgrounds, activity tends to start low in the morning, peak during mid-day, and taper off in the evening. Males consistently display slightly higher activity than females, with the most prominent differentiation in the "More than High School" category. Specifically, males in this group exhibit a sharp mid-day activity surge, contrasting with the more gradual rise observed in females. This data suggests that while daily routines may be somewhat influenced by education, gender differences in activity levels are consistently present across all educational tiers.


```{r,fig.width = 8,fig.asp =1.2,out.width = "90%"}
covar_add_accel %>% 
  pivot_longer(
    min1:min1440,
    names_to="minute",
    values_to="activity_level",
    names_prefix = "min"
  ) %>% 
  mutate(minute = as.numeric(minute) / 60) %>% 
  ggplot(aes(x=minute,y=activity_level,color=sex))+
  geom_point(alpha=0.7,size=1)+
  geom_smooth(se=FALSE)+
  facet_grid(education~.)+
  labs(
    title = "24-hour Activity Time Courses by Education Level and Sex",
    x = "Hours of Day (0-23)",
    y = "Activity Level",
    color = "Sex"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )+
  scale_x_continuous(
    breaks = seq(0, 23,2),
    labels = as.character(seq(0, 23,2)))
```

