---
title: "p8105_hw3_yw4251"
author: "You Wu"
date: 2023-10-10
output: github_document
---
# Problem 0
```{r}
library(tidyverse)
library(p8105.datasets)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1
```{r}
data("instacart")
instacart
```

## How many aisles are there, and which aisles are the most items ordered from?
```{r}
aisle_counts=
  instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n_obs=n()
  ) %>% 
  arrange(-n_obs)
```

There are `r nrow(aisle_counts)` aistles, and `r aisle_counts[1,1]` are the most items ordered from.

## Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
aisle_counts %>% 
  filter(n_obs>10000) %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n_obs)) %>% 
  ggplot(aes(x=aisle,y=n_obs))+
  geom_bar(stat = "identity",width=0.7)+
  labs(
    title = "Number of items ordered in each aisle",
    x = "Aisle Names",
    y = "Item Counts in Each Aisle",
    caption = "Data from the instacart")+
  scale_y_continuous(
    breaks = c(10000, 60000, 110000, 160000),
    labels = c("10000","60000", "110000","160000"))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r}
instacart %>% 
  filter(aisle%in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle,product_name) %>% 
  summarize(n_obs=n()) %>% 
  arrange(-n_obs) %>% 
  slice_head(n=3) %>% 
  mutate(rank = min_rank(desc(n_obs))) %>% 
  knitr::kable()
```

## Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
instacart %>% 
  filter(product_name%in%c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name,order_dow) %>% 
  summarize(mean_hour=mean(order_hour_of_day)) %>% 
  pivot_wider(names_from = order_dow,
              values_from=mean_hour) %>% 
  knitr::kable()
```


