---
title: "p8105_hw3_yw4251"
author: "You Wu"
date: 2023-10-10
output: github_document
---
# Problem 0
```{r}
library(tidyverse)
library(p8105.datasets)
library(ggridges)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .8,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1
```{r}
data("instacart")
instacart
```

## How many aisles are there, and which aisles are the most items ordered from?
```{r}
aisle_counts=
  instacart %>% 
  group_by(aisle) %>% 
  summarize(
    n_obs=n()
  ) %>% 
  arrange(-n_obs)
```

There are `r nrow(aisle_counts)` aistles, and `r aisle_counts[1,1]` are the most items ordered from.

## Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
aisle_counts %>% 
  filter(n_obs>10000) %>% 
  mutate(aisle = forcats::fct_reorder(aisle, n_obs)) %>% 
  ggplot(aes(x=aisle,y=n_obs))+
  geom_bar(stat = "identity",width=0.7)+
  labs(
    title = "Number of items ordered in each aisle",
    x = "Aisle Names",
    y = "Item Counts in Each Aisle",
    caption = "Data from the instacart")+
  scale_y_continuous(
    breaks = c(10000, 60000, 110000, 160000),
    labels = c("10000","60000", "110000","160000"))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

## Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r}
instacart %>% 
  filter(aisle%in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle,product_name) %>% 
  summarize(n_obs=n()) %>% 
  arrange(-n_obs) %>% 
  slice_head(n=3) %>% 
  mutate(rank = min_rank(desc(n_obs))) %>% 
  knitr::kable()
```

## Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
instacart %>% 
  filter(product_name%in%c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name,order_dow) %>% 
  summarize(mean_hour=mean(order_hour_of_day)) %>% 
  pivot_wider(names_from = order_dow,
              values_from=mean_hour) %>% 
  knitr::kable()
```


# Problem 2
```{r}
data("brfss_smart2010")
```

```{r}
cleaned_brfss=
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic=="Overall Health") %>% 
  mutate(response=factor(
    response,
    levels = c("Poor", "Fair", "Good", "Very good", "Excellent"),
    ordered = TRUE))
```

## In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
cleaned_brfss %>% 
  filter(year%in%c(2002,2010)) %>%
  group_by(year,locationabbr) %>% 
  summarize(n_loc=n_distinct(locationdesc)) %>% 
  filter(n_loc>=7) %>% 
  select(-n_loc) %>% 
  pivot_wider(
    names_from=year,
    values_from=locationabbr
  ) %>% 
  knitr::kable()
```

## Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).


```{r}
excel_cleaned_brfss=
  cleaned_brfss%>% 
  filter(response=="Excellent") %>% 
  group_by(year,locationabbr) %>% 
  summarize(avg_data_val=mean(data_value))

ggplot(excel_cleaned_brfss,aes(x=year,y=avg_data_val,color=locationabbr))+
  geom_line(size = 0.8,alpha=0.8)+
  theme_minimal()+
  theme(legend.text = element_text(size = 10))+
  labs(
    title = "Spaghetti Chart of Average Value over Time within States",
    x = "Year",
    y = "Average Data Values Across locations within State",
    caption = "Data from the Excellent Cleaned Brfss") +
  theme(
    plot.title = element_text(hjust = 0.5)
  )
  
```

## Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
ny_cleaned_brfss=
  cleaned_brfss %>% 
  filter(year%in%c(2006,2010)&locationabbr=="NY")


ggplot(ny_cleaned_brfss,aes(x=data_value,y=response,fill=response))+
  geom_density_ridges(alpha=0.3,scale=0.7,bins=40)+
  facet_grid(~year)+
  labs(
    title = "Distribution of Response Values in NY State (2006 vs 2010)",
    x = "Data Value",
    y = "Ordered Response Level",
    caption = "Data from the NY Cleaned Brfss"
    ) +
  theme_ridges(font_size = 14)+
  theme(
    legend.position = "none",
    axis.title.x = element_text(hjust = 0.5),
    axis.title.y = element_text(hjust = 0.5)
  )

```

# Problem 3

```{r}
covar_df=read_csv("data/nhanes_covar.csv",skip=4) %>% 
  janitor::clean_names() %>% 
  drop_na() %>%
  filter(age>=21) %>% 
  mutate(
    sex=as.factor(sex),
    sex=recode(sex, "1"="male", "2"="female"),
    education=factor(
      education,
      levels = c("1", "2", "3"),
      labels = c("Less than high school", "High school equivalent", "More than high school"),
      ordered = TRUE)
         )
  
accel_df=read_csv("data/nhanes_accel.csv") %>% 
  janitor::clean_names()

covar_add_accel=left_join(covar_df,accel_df,by="seqn") %>% 
  drop_na()
  
```

## Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r}
covar_add_accel %>% 
  group_by(education,sex) %>% 
  summarize(count=n()) %>% 
  knitr::kable()
```

```{r}
ggplot(covar_add_accel,aes(x=sex,y=age,fill=sex))+
  geom_boxplot(alpha=0.5)+
  facet_grid(~education) + 
  labs(
    title = "Age Distributions by Sex Across Education Categories",
    x = "Time",
    y = "Age",
    fill = "Sex"
    )+
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```

## Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r}
covar_add_accel %>% 
  mutate(total_activity=rowSums(select(., min1:min1440))) %>%
  ggplot(aes(x=age,y=total_activity,color=sex))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)+
  facet_grid(~education)+
   labs(
    title = "Total Activity by Age, Sex, and Education Level",
    x = "Age",
    y = "Total Activity",
    color = "Sex"
  )+
  theme(
    plot.title = element_text(hjust = 0.5)
  )
```


## Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r,fig.width = 8,fig.asp =1.2,out.width = "90%"}
covar_add_accel %>% 
  pivot_longer(
    min1:min1440,
    names_to="minute",
    values_to="activity_level",
    names_prefix = "min"
  ) %>% 
  mutate(minute = as.numeric(minute) / 60) %>% 
  ggplot(aes(x=minute,y=activity_level,color=sex))+
  geom_point(alpha=0.7,size=1)+
  geom_smooth()+
  facet_grid(education~.)+
  labs(
    title = "24-hour Activity Time Courses by Education Level and Sex",
    x = "Hours of Day (0-23)",
    y = "Activity Level",
    color = "Sex"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5)
  )+
  scale_x_continuous(
    breaks = seq(0, 23,2),
    labels = as.character(seq(0, 23,2)))
```

